---
title: "Getting started with planscanR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with planscanR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Most chunks talk to a government portal, so they are shown but not executed
  # when the vignette is built.
  eval = FALSE
)
library(planscanR)
```

`planscanR` helps you find environmental-assessment records — Environmental
Impact Assessments (EIA), Strategic Environmental Assessments (SEA), and related
advice — across European government portals. It fetches records into a tidy
table, caches them on disk, and (optionally) downloads their PDF documents.

It is a pure-R fetcher: scoring records by topic relevance lives in the
companion package **planscanR.screen**. This vignette covers planscanR itself —
fetching and the offline cache.

## Fetch records

A single function, `get_assessments()`, retrieves records from a portal. You
choose the country with a two-letter code:

```{r}
records <- get_assessments("nl", limit = 20, download = FALSE)
records
```

Two arguments are worth knowing from the start:

* `limit` caps how many records come back. The portals hold thousands of
  records, so always start small while you are exploring.
* `download` controls whether the actual PDF documents are fetched to your
  computer. Leave it `FALSE` until you have narrowed down which records you
  want — then the documents for those records are a quick second call away.

Whichever country you ask for, the result is a table with the same core
columns (`country`, `url`, `title`, `summary`, the document links, and so on),
so you can fetch from several countries and stack the results:

```{r}
nl <- get_assessments("nl", limit = 20, download = FALSE)
de <- get_assessments("de", limit = 20, download = FALSE)
both <- bind_results(nl, de)
```

To see which countries and search options are available, use:

```{r}
get_assessments_coverage()
```

## Working offline

Every record you fetch is saved to a small file on disk (a "sidecar") under the
cache root, which you can read back later without going to the portal at all:

```{r}
records <- index_cache(country = "nl")
```

This is handy for picking up where you left off, or for re-scoring an existing
set of records against new topics. The cache root resolves through
`cache_dir_default()`; `clear_cache()` wipes the downloaded files but leaves the
sidecars (and any derived artefacts) in place.

## Scoring records (optional)

`get_assessments()` accepts an optional `topic` (and `relevance_threshold`) to
score records during the fetch and gate PDF downloads on the score. The scoring
itself is delegated to the **planscanR.screen** package: install it to enable
`topic =`, and see its "Scoring records" vignette for the walkthrough. Without
it, omit `topic` and planscanR fetches without scoring — no Python required.

## Where to go next

* `?get_assessments` — every fetching option, including date and region filters.
* `?get_assessments_nl`, `?get_assessments_de`, `?get_assessments_at`,
  `?get_assessments_dk`, `?get_assessments_be`, `?get_assessments_ee` — the
  portal-specific details for each country.
* [**planscanR.screen**](https://barthoekstra.github.io/planscanR.screen/) —
  score records by topic relevance, classify them, and learn a selection model.
