---
title: "Getting started with planscanR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with planscanR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Most chunks talk to a government portal, so they are shown but not executed
  # when the vignette is built.
  eval = FALSE
)
library(planscanR)
```

`planscanR` helps you find environmental-assessment records — Environmental
Impact Assessments (EIA), Strategic Environmental Assessments (SEA), and related
advice — across European government portals. It fetches records into a tidy
table, caches them on disk, and (optionally) downloads their PDF documents.

It is a pure-R fetcher: scoring records by topic relevance lives in the
companion package **planscanR.screen**. This vignette covers planscanR itself —
fetching and the offline cache.

## Terminology

A few terms recur throughout planscanR and its documentation:

* **Record** — one result row: a single environmental-assessment case as
  returned by `get_assessments()`.
* **Assessment** — the case *type*: an Environmental Impact Assessment (EIA), a
  Strategic Environmental Assessment (SEA), or related advice.
* **Document** (or **attachment**) — a file belonging to a record, usually the
  assessment PDF, the decision, or a non-technical summary.
* **Offline metadata cache** — the on-disk store of fetched record metadata
  under the cache root, so records can be re-read without returning to the
  portal. Each record's metadata is one JSON file (a *sidecar*);
  `index_cache()` reads them all back.

## Fetch records

A single function, `get_assessments()`, retrieves records from a portal. You
choose the country with a two-letter code:

```{r}
records <- get_assessments("nl", limit = 20, download = FALSE)
records
```

Two arguments are worth knowing from the start:

* `limit` caps how many records come back. The portals hold thousands of
  records, so always start small while you are exploring.
* `download` controls whether the actual PDF documents are fetched to your
  computer. Leave it `FALSE` until you have narrowed down which records you
  want — then the documents for those records are a quick second call away.

Whichever country you ask for, the result is a table with the same core
columns (`country`, `url`, `title`, `summary`, the document links, and so on),
so you can fetch from several countries and stack the results:

```{r}
nl <- get_assessments("nl", limit = 20, download = FALSE)
de <- get_assessments("de", limit = 20, download = FALSE)
both <- bind_results(nl, de)
```

To see which countries and search options are available, use:

```{r}
get_assessments_coverage()
```

## Working offline

Every record you fetch is saved to the offline metadata cache under the cache
root, which you can read back later without going to the portal at all:

```{r}
records <- index_cache(country = "nl")
```

This is handy for picking up where you left off, or for re-scoring an existing
set of records against new topics. The cache root resolves through
`cache_dir_default()`; `clear_cache()` wipes the downloaded files but leaves the
cached record metadata (and any derived artefacts) in place.

## Scoring records (optional)

`get_assessments()` accepts an optional `topic` (and `relevance_threshold`) to
score records during the fetch and gate PDF downloads on the score:

```{r}
records <- get_assessments(
  "nl",
  topic = c(wind = "wind energy", solar = "solar energy"),
  relevance_threshold = 0.5,
  limit = 20,
  download = FALSE
)
# One score column per topic, plus the model that produced them.
records[c("title", "relevance_score_wind", "relevance_score_solar", "relevance_model")]
```

Each topic adds a `relevance_score_<name>` column. `relevance_threshold` only
*gates downloads* — it never drops records, so every record still appears in the
table and the offline metadata cache, and re-running with a different threshold
needs no network. The scoring itself is delegated to the **planscanR.screen**
package: install it to enable `topic =`, and see its "Scoring records" vignette
for the walkthrough. Without it, omit `topic` and planscanR fetches without
scoring — no Python required.

## Where to go next

* `?get_assessments` — every fetching option, including date and region filters.
* `?get_assessments_nl`, `?get_assessments_de`, `?get_assessments_fr`,
  `?get_assessments_at`, `?get_assessments_dk`, `?get_assessments_be`,
  `?get_assessments_ee`, `?get_assessments_fi`, `?get_assessments_bg`,
  `?get_assessments_cz`,
  `?get_assessments_hr`, `?get_assessments_gr`, `?get_assessments_is`,
  `?get_assessments_ie`, `?get_assessments_si`, `?get_assessments_pt`,
  `?get_assessments_gb`, `?get_assessments_it`, `?get_assessments_sk`,
  `?get_assessments_no` — the
  portal-specific details for each country.
* [**planscanR.screen**](https://barthoekstra.github.io/planscanR.screen/) —
  score records by topic relevance, classify them, and learn a selection model.
