---
title: "Supported sources"
output:
  rmarkdown::html_vignette:
    pandoc_args: ["--from=markdown+autolink_bare_uris+tex_math_single_backslash+alerts"]
vignette: >
  %\VignetteIndexEntry{Supported sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Every chunk in this vignette talks to a portal, so they are shown but
  # not executed at build time.
  eval = FALSE
)
library(planscanR)
```

`planscanR` is a thin layer over a handful of national environmental-assessment
portals (currently Netherlands, Germany, Austria, Denmark, Belgium (Flanders),
Estonia, and Bulgaria).
There is no shared API behind the scenes — each country gives us a
different mix of HTML pages, sitemaps, search endpoints, and undocumented JSON
handlers, and each one comes with its own quirks. This vignette describes what
each handler does, where it pulls data from, and what to expect (or watch out
for) when using it.

For the runtime equivalent — the same information in tabular form, including
the valid vocabularies for the search facets each handler accepts — call:

```{r}
get_assessments_coverage()
```

## A note on stability

> [!WARNING]
> None of these portals expose a contractually stable API. Some handlers parse
> HTML detail pages with `rvest` (NL, DE, EE), the rest sit on undocumented
> JSON or REST endpoints (AT, DK, BE). Any of the operators can change a CSS
> class,
> rename a field, or slip a login wall in front of an endpoint without
> warning, and the corresponding handler will then break — sometimes
> silently. Treat the results with the scepticism that a scraping pipeline
> deserves, and please file an issue when you spot differences.

> [!CAUTION]
> Because we are often not talking to real APIs, we have no rate-limit
> contract with these servers. The package throttles where we know it
> matters (NL is capped at ~1 request/second, DK and BE at 5), but a
> careless full-register crawl can still put real load on a small government
> portal. Use `limit` and `query` while exploring.

## Netherlands — `"nl"`

* **Portal:** Commissie m.e.r. adviezenregister
  ([commissiemer.nl/adviezen](https://www.commissiemer.nl/adviezen/))
* **Status:** supported — full records and document downloads.
* **Authentication:** none.
* **Coverage:** ~3,600 advisory records.
* **Throttle:** ~1 request/second by default
  (`getOption("planscanR.nl_throttle_rate")`).

URL enumeration is driven by the portal's own sitemap index (`wp-sitemap.xml`,
following the `advice-sitemap*` sub-sitemaps). Per-record metadata is parsed
from each detail page with `rvest` and persisted to a sidecar JSON, so
re-running a slice is essentially free once the cache is warm.

**Filter coverage.** Free-text `query`, `date_range` (matched against
`date_decision`), and `province` (substring match against
`competent_authority`) are honoured. The portal's taxonomy facets — `theme`,
`advice_type`, `status` — are driven by a client-side FacetWP layer that does
not yield to programmatic access; those arguments are accepted for forward
compatibility but currently emit a one-shot warning.

**Documents.** Detail pages group PDFs into two on-page sections, which are
exposed as separate list-columns (`attachment_urls_source` and the
deduplicated `attachment_urls` union). See `?get_assessments_nl` for the
section layout.

```{r}
get_assessments_nl(limit = 20, download = FALSE)
```

## Germany — `"de"`

* **Portal:** UVP-Verbund
  ([uvp-verbund.de](https://www.uvp-verbund.de/))
* **Status:** supported — full records and document downloads.
* **Authentication:** none.
* **Coverage:** ~24,289 records federated across all federal-state authorities
  (the full register).

The portal exposes no sitemap, no OAI-PMH, and no CSW endpoint. Enumeration
goes through the portal's Solr-backed full-text search (`/freitextsuche`).
When no `query` is supplied the handler sends the wildcard `q=*:*`, which
paginates through the full register (~2,429 pages of 10 records each). The
`toggle_procedure=` parameter is explicitly set to an empty value so the
portal does not silently restrict results to currently-running and
last-year-modified procedures.

**Filter coverage.** `query` is passed straight through as the server-side
`q` parameter (real full-text search, not a client-side substring match).
`date_range` matches against the detail-page "Zuletzt geändert" timestamp,
exposed as `date_decision`. `jurisdiction` is a substring match against the
federal-state partner (e.g. `"Bayern"`, `"Baden-Württemberg"`).

**Documents.** Detail pages group documents under open-ended `h4.title-font`
section headings. Each heading becomes its own
`attachment_urls_<slug>` / `local_path_<slug>` column; known headings get a
curated slug (`uvp_bericht`, `berichte`, `entscheidung`, `auslegung`,
`weitere`) and any unknown heading is auto-slugged from its title so a new
section type appears without a code change. See `?get_assessments_de` for the
full layout.

```{r}
get_assessments_de(query = "windenergie", limit = 20, download = FALSE)
```

## Austria — `"at"`

* **Portal:** Umweltbundesamt UVP-DB
  ([secure.umweltbundesamt.at/uvpdb/public](https://secure.umweltbundesamt.at/uvpdb/public))
* **Status:** supported (metadata-only) — documents sit behind a Keycloak
  login wall and are not retrievable by this version.
* **Authentication:** none for metadata.
* **Coverage:** ~500 procedures (the full register is small).

The public HTML pages of the portal sit behind a Keycloak login, but three
JSON service handlers are open: `mapsdata`, `mapsgeom`, and `vorhabenInfo`.
Enumeration is a single `mapsdata` call that returns the full index keyed by
Aktenzahl; per-record detail is one `vorhabenInfo` call per `v2id`. No
pagination, no CSRF, no session required.

**Filter coverage.** `query` is a case-insensitive substring match on
`title` + `summary`. `date_range` is matched against the record's `year`
(treated as a Jan 1 – Dec 31 span) — `date_decision` is always `NA`, because
the anonymous service handlers do not expose a decision or last-modified
timestamp. `jurisdiction` is a substring match against the comma-joined
Austrian federal-state list (so `"Wien"` works, `"Bayern"` will not).

**Documents.** Not available to anonymous callers. Every record returns
`attachment_urls = character(0)`. The `download` argument is accepted for
API symmetry but has no effect. If you need attachments, point
`get_assessments(..., discover = TRUE)` at a web-search backend — see
[`?discover_attachments`].

```{r}
get_assessments_at(query = "Windpark", limit = 20, download = FALSE)
```

## Denmark — `"dk"`

* **Portal:** Danmarks Miljøportal EA-Hub
  ([eahub.miljoeportal.dk](https://eahub.miljoeportal.dk/))
* **Status:** supported (metadata-only) — full metadata and polygon geometry
  are fetched; document downloads are a future addition.
* **Authentication:** none.
* **Coverage:** ~2,700 records (both EIA, "miljøvurdering af projekter", and
  SEA, "miljøvurdering af planer").
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.dk_throttle_rate")`), because the geometry fetch is
  one tiny GET per record-with-geometry.

EA-Hub is a Vue SPA sitting on a public REST API (Swagger at
`/api/swagger/v1/swagger.json`). One `POST /assessments/search` call returns
the entire register — each row already carries title, year range, status,
authorities, EIA-Directive Annex I/II categories, plan types/categories, and
a `hasGeometry` flag — so no detail call is needed during the scan phase.

**Geometry.** Records with `hasGeometry == TRUE` carry a polygon (typically a
MULTIPOLYGON in EPSG:25832 / ETRS89-UTM32N). When `write_sidecar = TRUE`, the
geometry is fetched from `GET /assessments/{id}/geometry` and saved next to
the sidecar as `<document_id>.geometry.geojson`; the sidecar carries a
`geometry_path` and `geometry_crs` for downstream consumption with `sf`.

**Filter coverage.** `query` is forwarded to the API's server-side `freeText`.
`assessment_type` (`"All"`, `"Plans"`, `"Project"`) is honoured. `date_range`
is matched client-side against each record's `fromYear`/`toYear` (treated as
Jan 1 – Dec 31 spans); `date_decision` is always `NA` because EA-Hub exposes
only year fields, no decision timestamp.

**Documents.** EA-Hub exposes PDFs at public Azure blob URLs reachable via
`GET /assessments/{id}/documents/{docId}/links`, but resolving them costs an
extra HTTP call per document. The current handler is scan + classify only;
a future release will add the download phase.

```{r}
get_assessments_dk(query = "vindmølle", limit = 20, download = FALSE)
```

## Belgium (Flanders) — `"be"`

* **Portal:** Departement Omgeving MER-register
  ([merregister.omgeving.vlaanderen.be](https://merregister.omgeving.vlaanderen.be/))
* **Status:** supported — full metadata, polygon geometry, and document
  downloads.
* **Authentication:** none.
* **Coverage:** ~3,000 Project-MER dossiers (project-level EIA) and dossier
  MER-plicht ontheffingsaanvragen (exemption requests). Plan-MER (SEA) lives
  in a separate Flemish register and is out of scope for this handler.
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.be_throttle_rate")`).

The portal is a Vue SPA backed by a public REST API at
`https://dmvb.omgeving.vlaanderen.be/api/v1/`. The SPA reads its backend host
from `/rest/configuratie` and then paginates `/api/v1/dossier` to enumerate
the register (page size capped server-side at 25). The full per-record
payload — coordinator, expertise domains, document list, geometry — is one
`/api/v1/dossier/{nummer}` call.

**Geometry.** Every detail record carries a `locatie` field in
GeoJSON-style (typically MULTIPOLYGON) directly inline — no separate
geometry call. Coordinates are in **EPSG:31370** (Belgian Lambert 72). When
`write_sidecar = TRUE`, the geometry is saved next to the sidecar as
`<document_id>.geometry.geojson`; the sidecar carries `geometry_path` and
`geometry_crs` for downstream consumption with `sf`.

**Filter coverage.** `query` is a case-insensitive substring match on
`title` + `document_id`. `niscode` (5-digit municipality code) and `nummer`
(dossier ID, e.g. `"PR4037"`) are forwarded server-side; the API ignores
anything else. `dossier_type` (`"PROJECT_MER"` /
`"VERZOEK_TOT_ONTHEFFING"`) is applied client-side. `date_range` matches
against `date_published`, derived from the earliest document creation date
on the record; `date_decision` is always `NA` because the API exposes no
separate decision timestamp.

**Documents.** Direct, anonymous download URLs are emitted for every
`documenten[]` entry. Documents are grouped by their portal `type`
(`Aanmelding`, `Ontheffingsaanvraag`, `Verslag toekenning ontheffing`,
...); each type becomes its own `attachment_urls_<slug>` /
`local_path_<slug>` column. The deduplicated union is at
`attachment_urls` / `local_path` as the schema requires.

```{r}
get_assessments_be(query = "wind", limit = 20, download = FALSE)
```

## Estonia — `"ee"`

* **Portal:** Keskkonnaamet KOTKAS (KMH-register and KSH-register)
  ([kotkas.envir.ee](https://kotkas.envir.ee/))
* **Status:** supported — full metadata, polygon geometry, and document
  downloads, across **both** registers.
* **Authentication:** none.
* **Coverage:** ~500 EIA records (KMH, *Keskkonnamõju hindamine*) and ~750
  SEA records (KSH, *Keskkonnamõju strateegiline hindamine*) at the time of
  writing. The handler merges both registers into one result tibble and
  tags each row via the `assessment_type` column (`"EIA"` for KMH, `"SEA"`
  for KSH). `document_id` is prefixed `"KMH-"` / `"KSH-"` so the two
  registers never collide on disk.
* **Throttle:** 2 requests/second by default
  (`getOption("planscanR.ee_throttle_rate")`) — the portal pushes back with a
  300 s retry-backoff at 5 req/s under sustained load.

KOTKAS is a server-rendered jQuery/Bootstrap portal (not a SPA). Each
register's index paginates via a numeric `qs=` offset (page size = 40,
server-controlled); detail pages live at `/kmh/kmh_view?kmh_id=<id>` and
`/kmh/ksh_view?ksh_id=<id>` respectively. Every field a downstream
classifier needs — full title, narrative summary, developer/proponent,
decider/competent authority, KOV municipality, geometry, attachment URLs —
is on the detail page.

**Geometry.** Every detail record carries its activity area as an inline
GeoJSON geometry, embedded in a hidden form input
(`#activity_area_geojson`). Coordinates are in **EPSG:3301** (Estonian
Coordinate System of 1997 / L-EST97). When `write_sidecar = TRUE`, the
geometry is saved next to the sidecar as `<document_id>.geometry.geojson`;
the sidecar carries `geometry_path` and `geometry_crs` for downstream
consumption with `sf`.

**Filter coverage.** `query` is forwarded server-side as
`s__search_keyword` (matches title / code / related-person).
`assessment_type` (`"All"`, `"EIA"`, `"SEA"`) decides which register(s) to
crawl. `proceeding_status` (`"INITIATED"`, `"ONGOING"`, `"SUSPENDED"`,
`"FINISHED"`), `activity_area` (maakond code, e.g. `"0037"` = Harju), and
`activity` (sector code, e.g. `"1300"` = energy) are all forwarded
server-side. `date_range` is matched client-side against `date_published`
(the portal's *Algatamise kpv* / initiation date); `date_decision` is
always `NA` because the portal does not expose a separate decision
timestamp on the detail page (only per-document dates inside the
*Dokumendid* table).

**Documents.** Detail pages expose a *Dokumendid* table whose rows have
direct, anonymous download URLs (`/kmh/<kmh|ksh>_file_download?...`).
Documents are grouped by their portal *Liik* (type) — common ones include
*Algatamise otsus* (initiation decision), *Programm* (assessment
programme), *Programmi otsus* (programme decision), *Aruanne* (report),
*Aruande otsus* (report decision). Each type becomes its own
`attachment_urls_<slug>` / `local_path_<slug>` column. The deduplicated
union is at `attachment_urls` / `local_path` as the schema requires.

```{r}
# Both registers, wind-themed slice
get_assessments_ee(query = "tuulepark", limit = 20, download = FALSE)

# SEA only
get_assessments_ee(assessment_type = "SEA", limit = 20, download = FALSE)
```

## Bulgaria — `"bg"`

* **Portal:** Ministry of Environment and Water (МОСВ) public registers —
  ОВОС-register (EIA) and ЕО-register (SEA)
  ([registers.moew.government.bg](https://registers.moew.government.bg/))
* **Status:** supported — full metadata and anonymous document downloads,
  across **both** registers. No geometry.
* **Authentication:** none.
* **Coverage:** ~35,000 EIA records (ОВОС, *Оценка на въздействието върху
  околната среда*) and ~10,000 SEA records (ЕО, *Екологична оценка*) at the
  time of writing. The handler merges both registers into one result tibble
  and tags each row via the `assessment_type` column (`"EIA"` for ОВОС,
  `"SEA"` for ЕО). `document_id` is prefixed `"OVOS-"` / `"EO-"` so the two
  registers never collide on disk. Because the registers are large, pass
  `limit` and/or `query` while exploring.
* **Throttle:** 2 requests/second by default
  (`getOption("planscanR.bg_throttle_rate")`) — detail fetches are slow
  (~4 s server-side) and the host is a small government server.

The registers are server-rendered ASP.NET MVC pages (no SPA, no viewstate).
Each register's listing paginates via a numeric `offset` plus a `limit`
query parameter (`?offset=<n>&limit=<k>`); the page text *Намерени <N>
досиета.* ("Found N dossiers") is the authoritative total. Detail pages live
at `/ovos/lot/<id>` and `/eo/lot/<id>` and are scraped with `rvest` from a
single nested row-group `table.table-lot` (each row carries a label `<th>`
and a value `<td>`).

**Geometry.** None. The МОСВ registers expose no coordinates or GeoJSON;
location is administrative text only and is composed into `jurisdiction`
(Област / Община / Населено място). No `geometry_path` / `geometry_crs`
columns are emitted.

**Filter coverage.** `query` is forwarded server-side as `projectName`
(substring match on the project/plan name). `assessment_type` (`"All"`,
`"EIA"`, `"SEA"`) decides which register(s) to crawl. `date_range` is matched
client-side against `date_published` (the dossier submission date);
`date_decision` is the termination-decision date (*Дата на решението за
прекратяване*) when present, else `NA`. The portal also accepts
`contractorNames` and region/authority/procedure/date filters server-side;
these are not first-class in v0.1.

**Documents.** Detail pages render document links inline inside the lot
table's value cells, with direct anonymous download URLs of the form
`/ovos/file?fileKey=<uuid>&fileName=<name>` (and `/eo/file?...`). The
`fileName` query parameter is required by the server (it 500s without it), so
the handler captures the full href verbatim rather than reconstructing it.
Documents are typed only by the table-row label they sit under (Уведомление,
Описание, Писмо, …); each label becomes its own `attachment_urls_<slug>` /
`local_path_<slug>` column (the Bulgarian label is transliterated to ASCII).
The deduplicated union is at `attachment_urls` / `local_path` as the schema
requires.

```{r}
# Both registers, wind-themed slice
get_assessments_bg(query = "вятър", limit = 20, download = FALSE)

# SEA only
get_assessments_bg(assessment_type = "SEA", limit = 20, download = FALSE)
```

## Czech Republic — `"cz"`

* **Portal:** CENIA *Informační systém EIA/SEA* — the EIA register *Záměry
  na území ČR* (`eia100_cr`) and the SEA register *Posuzování koncepcí*
  (`SEA100_koncepce`)
  ([portal.cenia.cz/eiasea](https://portal.cenia.cz/eiasea/))
* **Status:** supported — full metadata and anonymous document downloads,
  across **both** in-scope registers. No geometry.
* **Authentication:** none.
* **Scope (domestic CZ only):** the handler crawls **only** the two domestic
  registers above. The portal also hosts cross-border / foreign registers
  (`eia100_mimo_cr` — projects outside CZ, `sea100_mezistatni` — cross-border
  SEA) and several special sub-registers (the sub-limit / priority-transport /
  large-project EIA registers `eia100_podlimitni` / `_pdz` / `_vzvp` /
  `eia244`, and the territorial-planning SEA registers `sea100_pur*` /
  `zur*` / `up*`); these are **never** enumerated. Ministry-coded records
  (`EIA_MZP*` / `SEA_MZP*`) *inside* the two domestic registers are domestic
  projects/plans assessed by the Ministry and stay in scope.
* **Coverage:** ~22,780 EIA záměry and ~640 SEA koncepce at the time of
  writing. Both registers merge into one result tibble and tag each row via
  the `assessment_type` column (`"EIA"` / `"SEA"`). `document_id` is the
  register-namespaced detail code (e.g. `"EIA_JHC1237"`, `"SEA_HKK015K"`), so
  the two never collide on disk. Because the EIA register is large, pass
  `limit` while exploring.
* **Throttle:** 2 requests/second by default
  (`getOption("planscanR.cz_throttle_rate")`) — the portal is a single Tomcat
  instance and a full crawl is large.
* **Language:** record content is Czech (ISO-639-1 `cs`; the country code
  `cz` differs from the language code). `lang=en` only flips UI chrome, so the
  handler fetches with `lang=cs`.

The registers are server-rendered JSP pages (Apache Tomcat — not a SPA).
Each listing paginates via a **1-based** `?p=<n>` query (10 records/page);
out-of-range pages are clamped to the last page by the server, so the handler
stops paginating once a page contributes no detail codes it has not already
seen. Detail pages live at `/eiasea/detail/EIA_<CODE>` and
`/eiasea/detail/SEA_<CODE>` and are scraped with `rvest` from a single
`table.detail` of label/value rows interspersed with bold process-stage
heading rows.

**Geometry.** None. The CENIA detail pages expose no coordinates or GeoJSON;
location is administrative text only and is composed into `jurisdiction`
(Kraj / Okres / Obec / Katastr). No `geometry_path` / `geometry_crs` columns
are emitted.

**Filter coverage.** `assessment_type` (`"All"`, `"EIA"`, `"SEA"`) decides
which register(s) to crawl. `date_range` is matched client-side against
`date_published` — the EIA last-modified date (*Datum a čas posledních
úprav*) or the SEA publication date (*Datum zveřejnění*); `date_decision` is
always `NA` because the portal exposes no single clean decision-date field.
Dates are parsed from the Java `Date.toString()` form (e.g.
`Thu Jun 04 07:28:50 CEST 2026`).

**Documents.** Detail pages render document links inline inside the detail
table's value cells, with direct anonymous download URLs of the form
`/eiasea/download/<token>/<file>`. Both the base64-ish token and the trailing
human filename are captured verbatim from the rendered href. Documents are
grouped by process-stage heading (OZNÁMENÍ, ZJIŠŤOVACÍ ŘÍZENÍ, DOKUMENTACE,
POSUDEK, VEŘEJNÉ PROJEDNÁNÍ, STANOVISKO for EIA; the oznámení / vyhodnocení /
návrh koncepce / stanovisko / schválená koncepce fields for SEA); each stage
becomes its own `attachment_urls_<slug>` / `local_path_<slug>` column (the
Czech label is transliterated to ASCII). The deduplicated union is at
`attachment_urls` / `local_path` as the schema requires. Note that some
attachments are very large ZIP bundles (e.g. a 79 MB `oznameni.zip`); the
`max_file_size_mb` cap skips oversized files rather than fetching them.

```{r}
# Both registers, first few records
get_assessments_cz(limit = 5, download = FALSE)

# SEA only
get_assessments_cz(assessment_type = "SEA", limit = 5, download = FALSE)
```

## Adding a new source

Adding a handler for another national portal is a matter of writing one
`get_assessments_<cc>()` that returns a tibble with the planscanR required
columns (`country`, `source_portal`, `document_id`, `url`, `retrieved_at`,
`attachment_urls`, `local_path`) and wiring it into the dispatcher in
`get_assessments()`. If you have a portal in mind, open an issue describing
what it exposes (sitemap? search? open JSON?) and we can sketch out a handler.
