---
title: "Supported sources"
output:
  rmarkdown::html_vignette:
    pandoc_args: ["--from=markdown+autolink_bare_uris+tex_math_single_backslash+alerts"]
vignette: >
  %\VignetteIndexEntry{Supported sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Every chunk in this vignette talks to a portal, so they are shown but
  # not executed at build time.
  eval = FALSE
)
library(planscanR)
```

`planscanR` is a thin layer over a handful of national environmental-assessment
portals (currently Netherlands, Germany, France, Austria, Denmark, Belgium
(Flanders), Estonia, Finland, Bulgaria, the Czech Republic, Croatia, Greece,
Iceland, Ireland, and Slovenia).
There is no shared API behind the scenes — each country gives us a
different mix of HTML pages, sitemaps, search endpoints, and undocumented JSON
handlers, and each one comes with its own quirks. This vignette describes what
each handler does, where it pulls data from, and what to expect (or watch out
for) when using it.

For the runtime equivalent — the same information in tabular form, including
the valid vocabularies for the search facets each handler accepts — call:

```{r}
get_assessments_coverage()
```

## A note on stability

> [!WARNING]
> None of these portals expose a contractually stable API. Some handlers parse
> HTML detail pages with `rvest` (NL, DE, EE), the rest sit on undocumented
> JSON or REST endpoints (AT, DK, BE). Any of the operators can change a CSS
> class,
> rename a field, or slip a login wall in front of an endpoint without
> warning, and the corresponding handler will then break — sometimes
> silently. Treat the results with the scepticism that a scraping pipeline
> deserves, and please file an issue when you spot differences.

> [!CAUTION]
> Because we are often not talking to real APIs, we have no rate-limit
> contract with these servers. The package throttles where we know it
> matters (NL is capped at ~1 request/second, DK and BE at 5), but a
> careless full-register crawl can still put real load on a small government
> portal. Use `limit` and `query` while exploring.

## Netherlands — `"nl"`

* **Portal:** Commissie m.e.r. adviezenregister
  ([commissiemer.nl/adviezen](https://www.commissiemer.nl/adviezen/))
* **Status:** supported — full records and document downloads.
* **Authentication:** none.
* **Coverage:** ~3,600 advisory records.
* **Throttle:** ~1 request/second by default
  (`getOption("planscanR.nl_throttle_rate")`).

URL enumeration is driven by the portal's own sitemap index (`wp-sitemap.xml`,
following the `advice-sitemap*` sub-sitemaps). Per-record metadata is parsed
from each detail page with `rvest` and persisted to a sidecar JSON, so
re-running a slice is essentially free once the cache is warm.

**Filter coverage.** Free-text `query`, `date_range` (matched against
`date_decision`), and `province` (substring match against
`competent_authority`) are honoured. The portal's taxonomy facets — `theme`,
`advice_type`, `status` — are driven by a client-side FacetWP layer that does
not yield to programmatic access; those arguments are accepted for forward
compatibility but currently emit a one-shot warning.

**Documents.** Detail pages group PDFs into two on-page sections, which are
exposed as separate list-columns (`attachment_urls_source` and the
deduplicated `attachment_urls` union). See `?get_assessments_nl` for the
section layout.

```{r}
get_assessments_nl(limit = 20, download = FALSE)
```

## Germany — `"de"`

* **Portal:** UVP-Verbund
  ([uvp-verbund.de](https://www.uvp-verbund.de/))
* **Status:** supported — full records and document downloads.
* **Authentication:** none.
* **Coverage:** ~24,289 records federated across all federal-state authorities
  (the full register).

The portal exposes no sitemap, no OAI-PMH, and no CSW endpoint. Enumeration
goes through the portal's Solr-backed full-text search (`/freitextsuche`).
When no `query` is supplied the handler sends the wildcard `q=*:*`, which
paginates through the full register (~2,429 pages of 10 records each). The
`toggle_procedure=` parameter is explicitly set to an empty value so the
portal does not silently restrict results to currently-running and
last-year-modified procedures.

**Filter coverage.** `query` is passed straight through as the server-side
`q` parameter (real full-text search, not a client-side substring match).
`date_range` matches against the detail-page "Zuletzt geändert" timestamp,
exposed as `date_decision`. `jurisdiction` is a substring match against the
federal-state partner (e.g. `"Bayern"`, `"Baden-Württemberg"`).

**Documents.** Detail pages group documents under open-ended `h4.title-font`
section headings. Each heading becomes its own
`attachment_urls_<slug>` / `local_path_<slug>` column; known headings get a
curated slug (`uvp_bericht`, `berichte`, `entscheidung`, `auslegung`,
`weitere`) and any unknown heading is auto-slugged from its title so a new
section type appears without a code change. See `?get_assessments_de` for the
full layout.

```{r}
get_assessments_de(query = "windenergie", limit = 20, download = FALSE)
```

## France — `"fr"`

* **Portal:** Projets-Environnement
  ([projets-environnement.gouv.fr](https://www.projets-environnement.gouv.fr/))
* **Status:** supported — full records, WGS84 geometry, and document downloads.
* **Authentication:** none.
* **Coverage:** ~5,483 records (the full diffusion dataset).

Unlike the SPA-behind-a-private-API portals elsewhere in the family, this one
is a public **OpenDataSoft Explore API v2.1** platform — a documented,
anonymous REST+JSON service. The whole register is one flat dataset
(`projets-environnement-diffusion`), with every field inline in each record,
so there is no separate detail call. The handler enumerates via the export
endpoint (`/exports/json?limit=-1`), which has no offset cap and returns the
full filtered set in one call. The stable business key `recordsid` is used as
`document_id`.

**Throttle.** 5 requests per second by default
(`getOption("planscanR.fr_throttle_rate")`). The enumeration is a single
export call; the throttle mainly paces per-attachment downloads.

**Filter coverage.** All filters are server-side via ODSQL `where` clauses:
`query` becomes `search("<query>")`; `theme` (`dc_subject_theme`),
`native_type` (`dc_type`) and `status` (`vp_status`) become equality clauses;
`date_range` becomes a `dc_date` window. See `?get_assessments_coverage` for
the `theme` / `status` / `native_type` vocabularies.

**Documents.** Attachments come from a fixed set of typed `dc_relation_*`
fields, each mapped to a curated slug: `attachment_urls_etude_impact` (the
étude d'impact PDF), `_resume_non_technique` (résumé non technique),
`_avis_ae` (avis de l'Autorité environnementale), `_reponse_avis_ae`,
`_dossier` (the `*_DCZIP.zip`), plus a few others. Only real document URLs on
the `sicodei.projets-environnement.gouv.fr` blob host (or `.pdf`/`.zip` links)
become attachments — external préfecture web pages are kept only as extras.

**Geometry.** About 1,472 records carry a `localisation` GeoJSON Feature
(typically a `MultiPolygon`). OpenDataSoft always serves WGS84, so each is
saved as a sibling `<document_id>.geometry.geojson` with
`geometry_crs = "EPSG:4326"`.

```{r}
get_assessments_fr(query = "éolien", limit = 20, download = FALSE)
```

## Austria — `"at"`

* **Portal:** Umweltbundesamt UVP-DB
  ([secure.umweltbundesamt.at/uvpdb/public](https://secure.umweltbundesamt.at/uvpdb/public))
* **Status:** supported (metadata-only) — documents sit behind a Keycloak
  login wall and are not retrievable by this version.
* **Authentication:** none for metadata.
* **Coverage:** ~500 procedures (the full register is small).

The public HTML pages of the portal sit behind a Keycloak login, but three
JSON service handlers are open: `mapsdata`, `mapsgeom`, and `vorhabenInfo`.
Enumeration is a single `mapsdata` call that returns the full index keyed by
Aktenzahl; per-record detail is one `vorhabenInfo` call per `v2id`. No
pagination, no CSRF, no session required.

**Filter coverage.** `query` is a case-insensitive substring match on
`title` + `summary`. `date_range` is matched against the record's `year`
(treated as a Jan 1 – Dec 31 span) — `date_decision` is always `NA`, because
the anonymous service handlers do not expose a decision or last-modified
timestamp. `jurisdiction` is a substring match against the comma-joined
Austrian federal-state list (so `"Wien"` works, `"Bayern"` will not).

**Documents.** Not available to anonymous callers. Every record returns
`attachment_urls = character(0)`. The `download` argument is accepted for
API symmetry but has no effect. If you need attachments, point
`get_assessments(..., discover = TRUE)` at a web-search backend — see
[`?discover_attachments`].

```{r}
get_assessments_at(query = "Windpark", limit = 20, download = FALSE)
```

## Denmark — `"dk"`

* **Portal:** Danmarks Miljøportal EA-Hub
  ([eahub.miljoeportal.dk](https://eahub.miljoeportal.dk/))
* **Status:** supported (metadata-only) — full metadata and polygon geometry
  are fetched; document downloads are a future addition.
* **Authentication:** none.
* **Coverage:** ~2,700 records (both EIA, "miljøvurdering af projekter", and
  SEA, "miljøvurdering af planer").
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.dk_throttle_rate")`), because the geometry fetch is
  one tiny GET per record-with-geometry.

EA-Hub is a Vue SPA sitting on a public REST API (Swagger at
`/api/swagger/v1/swagger.json`). One `POST /assessments/search` call returns
the entire register — each row already carries title, year range, status,
authorities, EIA-Directive Annex I/II categories, plan types/categories, and
a `hasGeometry` flag — so no detail call is needed during the scan phase.

**Geometry.** Records with `hasGeometry == TRUE` carry a polygon (typically a
MULTIPOLYGON in EPSG:25832 / ETRS89-UTM32N). When `write_sidecar = TRUE`, the
geometry is fetched from `GET /assessments/{id}/geometry` and saved next to
the sidecar as `<document_id>.geometry.geojson`; the sidecar carries a
`geometry_path` and `geometry_crs` for downstream consumption with `sf`.

**Filter coverage.** `query` is forwarded to the API's server-side `freeText`.
`assessment_type` (`"All"`, `"Plans"`, `"Project"`) is honoured. `date_range`
is matched client-side against each record's `fromYear`/`toYear` (treated as
Jan 1 – Dec 31 spans); `date_decision` is always `NA` because EA-Hub exposes
only year fields, no decision timestamp.

**Documents.** EA-Hub exposes PDFs at public Azure blob URLs reachable via
`GET /assessments/{id}/documents/{docId}/links`, but resolving them costs an
extra HTTP call per document. The current handler is scan + classify only;
a future release will add the download phase.

```{r}
get_assessments_dk(query = "vindmølle", limit = 20, download = FALSE)
```

## Belgium (Flanders) — `"be"`

* **Portal:** Departement Omgeving MER-register
  ([merregister.omgeving.vlaanderen.be](https://merregister.omgeving.vlaanderen.be/))
* **Status:** supported — full metadata, polygon geometry, and document
  downloads.
* **Authentication:** none.
* **Coverage:** ~3,000 Project-MER dossiers (project-level EIA) and dossier
  MER-plicht ontheffingsaanvragen (exemption requests). Plan-MER (SEA) lives
  in a separate Flemish register and is out of scope for this handler.
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.be_throttle_rate")`).

The portal is a Vue SPA backed by a public REST API at
`https://dmvb.omgeving.vlaanderen.be/api/v1/`. The SPA reads its backend host
from `/rest/configuratie` and then paginates `/api/v1/dossier` to enumerate
the register (page size capped server-side at 25). The full per-record
payload — coordinator, expertise domains, document list, geometry — is one
`/api/v1/dossier/{nummer}` call.

**Geometry.** Every detail record carries a `locatie` field in
GeoJSON-style (typically MULTIPOLYGON) directly inline — no separate
geometry call. Coordinates are in **EPSG:31370** (Belgian Lambert 72). When
`write_sidecar = TRUE`, the geometry is saved next to the sidecar as
`<document_id>.geometry.geojson`; the sidecar carries `geometry_path` and
`geometry_crs` for downstream consumption with `sf`.

**Filter coverage.** `query` is a case-insensitive substring match on
`title` + `document_id`. `niscode` (5-digit municipality code) and `nummer`
(dossier ID, e.g. `"PR4037"`) are forwarded server-side; the API ignores
anything else. `dossier_type` (`"PROJECT_MER"` /
`"VERZOEK_TOT_ONTHEFFING"`) is applied client-side. `date_range` matches
against `date_published`, derived from the earliest document creation date
on the record; `date_decision` is always `NA` because the API exposes no
separate decision timestamp.

**Documents.** Direct, anonymous download URLs are emitted for every
`documenten[]` entry. Documents are grouped by their portal `type`
(`Aanmelding`, `Ontheffingsaanvraag`, `Verslag toekenning ontheffing`,
...); each type becomes its own `attachment_urls_<slug>` /
`local_path_<slug>` column. The deduplicated union is at
`attachment_urls` / `local_path` as the schema requires.

```{r}
get_assessments_be(query = "wind", limit = 20, download = FALSE)
```

## Estonia — `"ee"`

* **Portal:** Keskkonnaamet KOTKAS (KMH-register and KSH-register)
  ([kotkas.envir.ee](https://kotkas.envir.ee/))
* **Status:** supported — full metadata, polygon geometry, and document
  downloads, across **both** registers.
* **Authentication:** none.
* **Coverage:** ~500 EIA records (KMH, *Keskkonnamõju hindamine*) and ~750
  SEA records (KSH, *Keskkonnamõju strateegiline hindamine*) at the time of
  writing. The handler merges both registers into one result tibble and
  tags each row via the `assessment_type` column (`"EIA"` for KMH, `"SEA"`
  for KSH). `document_id` is prefixed `"KMH-"` / `"KSH-"` so the two
  registers never collide on disk.
* **Throttle:** 2 requests/second by default
  (`getOption("planscanR.ee_throttle_rate")`) — the portal pushes back with a
  300 s retry-backoff at 5 req/s under sustained load.

KOTKAS is a server-rendered jQuery/Bootstrap portal (not a SPA). Each
register's index paginates via a numeric `qs=` offset (page size = 40,
server-controlled); detail pages live at `/kmh/kmh_view?kmh_id=<id>` and
`/kmh/ksh_view?ksh_id=<id>` respectively. Every field a downstream
classifier needs — full title, narrative summary, developer/proponent,
decider/competent authority, KOV municipality, geometry, attachment URLs —
is on the detail page.

**Geometry.** Every detail record carries its activity area as an inline
GeoJSON geometry, embedded in a hidden form input
(`#activity_area_geojson`). Coordinates are in **EPSG:3301** (Estonian
Coordinate System of 1997 / L-EST97). When `write_sidecar = TRUE`, the
geometry is saved next to the sidecar as `<document_id>.geometry.geojson`;
the sidecar carries `geometry_path` and `geometry_crs` for downstream
consumption with `sf`.

**Filter coverage.** `query` is forwarded server-side as
`s__search_keyword` (matches title / code / related-person).
`assessment_type` (`"All"`, `"EIA"`, `"SEA"`) decides which register(s) to
crawl. `proceeding_status` (`"INITIATED"`, `"ONGOING"`, `"SUSPENDED"`,
`"FINISHED"`), `activity_area` (maakond code, e.g. `"0037"` = Harju), and
`activity` (sector code, e.g. `"1300"` = energy) are all forwarded
server-side. `date_range` is matched client-side against `date_published`
(the portal's *Algatamise kpv* / initiation date); `date_decision` is
always `NA` because the portal does not expose a separate decision
timestamp on the detail page (only per-document dates inside the
*Dokumendid* table).

**Documents.** Detail pages expose a *Dokumendid* table whose rows have
direct, anonymous download URLs (`/kmh/<kmh|ksh>_file_download?...`).
Documents are grouped by their portal *Liik* (type) — common ones include
*Algatamise otsus* (initiation decision), *Programm* (assessment
programme), *Programmi otsus* (programme decision), *Aruanne* (report),
*Aruande otsus* (report decision). Each type becomes its own
`attachment_urls_<slug>` / `local_path_<slug>` column. The deduplicated
union is at `attachment_urls` / `local_path` as the schema requires.

```{r}
# Both registers, wind-themed slice
get_assessments_ee(query = "tuulepark", limit = 20, download = FALSE)

# SEA only
get_assessments_ee(assessment_type = "SEA", limit = 20, download = FALSE)
```

## Finland — `"fi"`

* **Portal:** ymparisto.fi, the national environmental-administration site
  ([ymparisto.fi](https://www.ymparisto.fi/))
* **Status:** supported — **EIA/YVA only**. The ymparisto.fi search index has
  no SOVA/SEA (strategic assessment) content type; `yva_project` is the only
  project type, so this handler delivers project-level EIA (*YVA*,
  *ympäristövaikutusten arviointi*) records only. There is no SEA path.
* **Authentication:** none.
* **Coverage:** ~1,334 YVA records at the time of writing.
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.fi_throttle_rate")`) — the ~1,334 per-record HTML
  landing-page GETs (needed to harvest attachment URLs) dominate a cold crawl.

This handler is a **hybrid**. ymparisto.fi is a Drupal + React site whose
site-search is backed by an Elasticsearch index exposed through a same-origin
proxy that passes raw ES Query DSL: `POST
https://www.ymparisto.fi/fi/app/search/query` with a JSON body. The handler
filters to `{"query":{"term":{"type":"yva_project"}}}` and paginates with ES
`from`/`size` against a stable id sort; the whole register sits under the
10,000 `max_result_window`. All record **metadata** (title, description,
publish time, project phase, the responsible ELY-keskus, municipality,
province, subject area) comes from the index. A Swedish `/sv/` index exists
but is ignored in v0.1.

**Geometry.** None — neither the index nor the landing page exposes
coordinates, so no geometry columns are emitted.

**Filter coverage.** `query` is forwarded server-side as an ES `match` over the
`content` + `title` fields. `assessment_type` accepts only `"All"` / `"EIA"`
(both crawl the single `yva_project` register; there is no `"SEA"`).
`date_range` is matched client-side against `date_published` (the `publishTime`
unix-seconds epoch); `date_decision` is always `NA` (no structured decision
timestamp in the index).

**Documents.** Attachment URLs are **not** in the Elasticsearch index — they
live only on the HTML landing page. For each kept record the handler GETs the
record's landing page and scrapes every `<a href>` under
`/sites/default/files/`, resolving them to absolute, anonymous
`https://www.ymparisto.fi/sites/default/files/documents/<file>.{pdf,doc,docx}`
URLs. Because the URLs come from the HTML, this detail fetch runs even when
`download = FALSE` (to populate `attachment_urls`); it is skipped only when a
sidecar already exists for the URL (sidecar-first). Documents are typed by
their **anchor text** (the page has no structured section markup) via a curated
keyword map with an auto-slug fallback: `arviointiohjelma` → `programme`,
`arviointiselostus` → `report`, `lausunto` → `statement`, `kuulutus` →
`notice`, `tiivistelmä` → `summary`. Each type becomes its own
`attachment_urls_<slug>` / `local_path_<slug>` column; the deduplicated union
is at `attachment_urls` / `local_path` as the schema requires.

```{r}
# Wind-themed slice (server-side full-text)
get_assessments_fi(query = "tuulivoima", limit = 20, download = FALSE)
```

## Bulgaria — `"bg"`

* **Portal:** Ministry of Environment and Water (МОСВ) public registers —
  ОВОС-register (EIA) and ЕО-register (SEA)
  ([registers.moew.government.bg](https://registers.moew.government.bg/))
* **Status:** supported — full metadata and anonymous document downloads,
  across **both** registers. No geometry.
* **Authentication:** none.
* **Coverage:** ~35,000 EIA records (ОВОС, *Оценка на въздействието върху
  околната среда*) and ~10,000 SEA records (ЕО, *Екологична оценка*) at the
  time of writing. The handler merges both registers into one result tibble
  and tags each row via the `assessment_type` column (`"EIA"` for ОВОС,
  `"SEA"` for ЕО). `document_id` is prefixed `"OVOS-"` / `"EO-"` so the two
  registers never collide on disk. Because the registers are large, pass
  `limit` and/or `query` while exploring.
* **Throttle:** 2 requests/second by default
  (`getOption("planscanR.bg_throttle_rate")`) — detail fetches are slow
  (~4 s server-side) and the host is a small government server.

The registers are server-rendered ASP.NET MVC pages (no SPA, no viewstate).
Each register's listing paginates via a numeric `offset` plus a `limit`
query parameter (`?offset=<n>&limit=<k>`); the page text *Намерени <N>
досиета.* ("Found N dossiers") is the authoritative total. Detail pages live
at `/ovos/lot/<id>` and `/eo/lot/<id>` and are scraped with `rvest` from a
single nested row-group `table.table-lot` (each row carries a label `<th>`
and a value `<td>`).

**Geometry.** None. The МОСВ registers expose no coordinates or GeoJSON;
location is administrative text only and is composed into `jurisdiction`
(Област / Община / Населено място). No `geometry_path` / `geometry_crs`
columns are emitted.

**Filter coverage.** `query` is forwarded server-side as `projectName`
(substring match on the project/plan name). `assessment_type` (`"All"`,
`"EIA"`, `"SEA"`) decides which register(s) to crawl. `date_range` is matched
client-side against `date_published` (the dossier submission date);
`date_decision` is the termination-decision date (*Дата на решението за
прекратяване*) when present, else `NA`. The portal also accepts
`contractorNames` and region/authority/procedure/date filters server-side;
these are not first-class in v0.1.

**Documents.** Detail pages render document links inline inside the lot
table's value cells, with direct anonymous download URLs of the form
`/ovos/file?fileKey=<uuid>&fileName=<name>` (and `/eo/file?...`). The
`fileName` query parameter is required by the server (it 500s without it), so
the handler captures the full href verbatim rather than reconstructing it.
Documents are typed only by the table-row label they sit under (Уведомление,
Описание, Писмо, …); each label becomes its own `attachment_urls_<slug>` /
`local_path_<slug>` column (the Bulgarian label is transliterated to ASCII).
The deduplicated union is at `attachment_urls` / `local_path` as the schema
requires.

```{r}
# Both registers, wind-themed slice
get_assessments_bg(query = "вятър", limit = 20, download = FALSE)

# SEA only
get_assessments_bg(assessment_type = "SEA", limit = 20, download = FALSE)
```

## Czech Republic — `"cz"`

* **Portal:** CENIA *Informační systém EIA/SEA* — the EIA register *Záměry
  na území ČR* (`eia100_cr`) and the SEA register *Posuzování koncepcí*
  (`SEA100_koncepce`)
  ([portal.cenia.cz/eiasea](https://portal.cenia.cz/eiasea/))
* **Status:** supported — full metadata and anonymous document downloads,
  across **both** in-scope registers. No geometry.
* **Authentication:** none.
* **Scope (domestic CZ only):** the handler crawls **only** the two domestic
  registers above. The portal also hosts cross-border / foreign registers
  (`eia100_mimo_cr` — projects outside CZ, `sea100_mezistatni` — cross-border
  SEA) and several special sub-registers (the sub-limit / priority-transport /
  large-project EIA registers `eia100_podlimitni` / `_pdz` / `_vzvp` /
  `eia244`, and the territorial-planning SEA registers `sea100_pur*` /
  `zur*` / `up*`); these are **never** enumerated. Ministry-coded records
  (`EIA_MZP*` / `SEA_MZP*`) *inside* the two domestic registers are domestic
  projects/plans assessed by the Ministry and stay in scope.
* **Coverage:** ~22,780 EIA záměry and ~640 SEA koncepce at the time of
  writing. Both registers merge into one result tibble and tag each row via
  the `assessment_type` column (`"EIA"` / `"SEA"`). `document_id` is the
  register-namespaced detail code (e.g. `"EIA_JHC1237"`, `"SEA_HKK015K"`), so
  the two never collide on disk. Because the EIA register is large, pass
  `limit` while exploring.
* **Throttle:** 2 requests/second by default
  (`getOption("planscanR.cz_throttle_rate")`) — the portal is a single Tomcat
  instance and a full crawl is large.
* **Language:** record content is Czech (ISO-639-1 `cs`; the country code
  `cz` differs from the language code). `lang=en` only flips UI chrome, so the
  handler fetches with `lang=cs`.

The registers are server-rendered JSP pages (Apache Tomcat — not a SPA).
Each listing paginates via a **1-based** `?p=<n>` query (10 records/page);
out-of-range pages are clamped to the last page by the server, so the handler
stops paginating once a page contributes no detail codes it has not already
seen. Detail pages live at `/eiasea/detail/EIA_<CODE>` and
`/eiasea/detail/SEA_<CODE>` and are scraped with `rvest` from a single
`table.detail` of label/value rows interspersed with bold process-stage
heading rows.

**Geometry.** None. The CENIA detail pages expose no coordinates or GeoJSON;
location is administrative text only and is composed into `jurisdiction`
(Kraj / Okres / Obec / Katastr). No `geometry_path` / `geometry_crs` columns
are emitted.

**Filter coverage.** `assessment_type` (`"All"`, `"EIA"`, `"SEA"`) decides
which register(s) to crawl. `date_range` is matched client-side against
`date_published` — the EIA last-modified date (*Datum a čas posledních
úprav*) or the SEA publication date (*Datum zveřejnění*); `date_decision` is
always `NA` because the portal exposes no single clean decision-date field.
Dates are parsed from the Java `Date.toString()` form (e.g.
`Thu Jun 04 07:28:50 CEST 2026`).

**Documents.** Detail pages render document links inline inside the detail
table's value cells, with direct anonymous download URLs of the form
`/eiasea/download/<token>/<file>`. Both the base64-ish token and the trailing
human filename are captured verbatim from the rendered href. Documents are
grouped by process-stage heading (OZNÁMENÍ, ZJIŠŤOVACÍ ŘÍZENÍ, DOKUMENTACE,
POSUDEK, VEŘEJNÉ PROJEDNÁNÍ, STANOVISKO for EIA; the oznámení / vyhodnocení /
návrh koncepce / stanovisko / schválená koncepce fields for SEA); each stage
becomes its own `attachment_urls_<slug>` / `local_path_<slug>` column (the
Czech label is transliterated to ASCII). The deduplicated union is at
`attachment_urls` / `local_path` as the schema requires. Note that some
attachments are very large ZIP bundles (e.g. a 79 MB `oznameni.zip`); the
`max_file_size_mb` cap skips oversized files rather than fetching them.

```{r}
# Both registers, first few records
get_assessments_cz(limit = 5, download = FALSE)

# SEA only
get_assessments_cz(assessment_type = "SEA", limit = 5, download = FALSE)
```

## Croatia — `"hr"`

* **Portal:** Ministry of Environment and Green Transition CMS pages —
  *Procjena utjecaja zahvata na okoliš* (PUO / project-level EIA) and
  *Strateška procjena utjecaja na okoliš* (SPUO / plan-level SEA)
  ([mzozt.gov.hr](https://mzozt.gov.hr/)). Pin the domain `mzozt.gov.hr`;
  it was renamed from `mingor` / `mingo` and the old links 30x-redirect to it.
* **Status:** supported — full metadata and anonymous document downloads,
  across **both** registers. No API. No geometry.
* **Authentication:** none.
* **No machine-readable register.** Croatia exposes no JSON API and no
  per-record detail endpoint. The "register" is a small set of
  server-rendered ASP.NET CMS pages; each procedure is an inlined
  `<li><strong>PROJECT TITLE</strong> <ul>...document links...</ul></li>`
  block. The handler fetches the master page(s) once and treats each block as
  one record — there is **no** pagination. PUO is one large master archive
  page (~550 procedures / ~2,500 documents); SPUO is two small pages
  (Ministry-competent and other-competent-body procedures).
* **Coverage:** both registers merge into one result tibble and tag each row
  via the `assessment_type` column (`"EIA"` / `"SEA"`). Because there is no
  native procedure id, `document_id` is a stable deterministic SHA-1 hash of
  the title (`HR-PUO-<hash>` / `HR-SPUO-<hash>`, with a cleanly-parseable year
  folded in when present), so the two never collide on disk. `url` is the
  master-page URL plus `#<document_id>`, giving each record a unique landing
  URL while still pointing a human at the right page.
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.hr_throttle_rate")`) — only ~3 HTML fetches plus N
  PDF downloads, so the throttle is politeness for the download phase.
* **Language:** record content is Croatian (ISO-639-1 `hr`).

**Geometry.** None. The CMS pages expose no coordinates or GeoJSON; spatial
information, where present, lives inside the PDFs. A county / *grad* is
heuristically pulled from the title into `jurisdiction` when trivially
present, else `NA`. No `geometry_path` / `geometry_crs` columns are emitted.

**Filter coverage.** `assessment_type` (`"All"`, `"EIA"`, `"SEA"`) decides
which register(s) to crawl. `query` is a **client-side** title substring match
(the CMS pages have no server-side search). `date_range` is matched
client-side against `date_published` — the earliest document date in the block
(every document anchor is prefixed `DD.MM.YYYY.`); `date_decision` is the
*rješenje* / *odluka* (decision) date when present, else `NA`.

**Documents.** Direct anonymous `.pdf` (sometimes `.zip`) download links carry
stable `data-fileid` attributes. Documents are grouped by the stage
sub-heading they sit under (PUO *informacija o zahtjevu* / *javni uvid* /
*nacrt rješenja* / *rješenje*); SPUO procedures often list their documents
flat under the title with no stage sub-heading, which fall under the
`document` slug. Known stage labels get a stable curated slug; anything else
is auto-slugged from the heading (Croatian diacritics transliterated to
ASCII). Each stage becomes its own `attachment_urls_<slug>` /
`local_path_<slug>` column; the deduplicated union is at `attachment_urls` /
`local_path` as the schema requires. The `href` is captured verbatim (with its
spaces and Croatian diacritics) and percent-encoded only for the network
request at download time.

```{r}
# Both registers, first few records
get_assessments_hr(limit = 5, download = FALSE)

# SEA only
get_assessments_hr(assessment_type = "SEA", download = FALSE)
```

## Greece — `"gr"`

* **Portal:** ΗΠΜ / EPRM (*Ηλεκτρονικό Περιβαλλοντικό Μητρώο* — Electronic
  Environmental Registry) of the Ministry of Environment & Energy (ΥΠΕΝ),
  served by a public JSON:API at
  [api.eprm.ypen.gr](https://api.eprm.ypen.gr/v1/license-decisions). The
  human-facing SPA lives at [eprm.ypen.gr](https://eprm.ypen.gr/) (its
  `robots.txt` disallows crawling the HTML, so the handler harvests only from
  the JSON:API and uses the SPA route purely as the canonical `url`).
* **Status:** supported, **with an important coverage caveat (decisions only)** —
  see below. Reflected in `get_assessments_coverage()$status` as
  `"supported (decisions-only; studies/SEA login-gated)"`.
* **Authentication:** none for the public registry (each request sends
  `Accept: application/json`; no cookie / CSRF).
* **AEPO decisions only (read this first).** The public registry exposes only
  **AEPO decisions** — *Αποφάσεις Έγκρισης Περιβαλλοντικών Όρων*, the
  regulatory **output** of the EIA (ΜΠΕ) process. The underlying **ΜΠΕ study
  files and all ΣΜΠΕ / SEA records sit behind the gov.gr login** on
  `platform.eprm.ypen.gr` and are **not** fetchable here. So each record is a
  decision (metadata + at most one decision PDF) — never the EIA study itself,
  and the SEA register is entirely out of scope. Plan for a decisions-only
  corpus.
* **Coverage:** ≈19,800 AEPO decisions. The listing endpoint
  (`GET /v1/license-decisions`) paginates JSON:API-style via `page[number]`
  (1-based) and `page[size]` (default 100); each list row is **already the full
  record**, so there is no separate detail call. `document_id` is the registry
  `id` (e.g. `"19895"`, globally unique in the register).
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.gr_throttle_rate")`) — enumeration is ≈200 list calls
  plus one download per attachment.
* **Language:** record content is Greek (ISO-639-1 `el`; note the country code
  `gr` differs from the language code `el`).

**Geometry.** A record's `project_location` is an array of `{lat, lon}` pairs.
When present, the first pair is written as a **Point** geometry next to the
sidecar as `<document_id>.geometry.geojson` (the family FeatureCollection
layout, GeoJSON-2008 `crs` member naming `urn:ogc:def:crs:EPSG::4326`). The
coordinates are already geographic **WGS84 (EPSG:4326)** — *not* the Greek Grid
(EPSG:2100) — so no reprojection happens; `geometry_crs` is `"EPSG:4326"`.
Records with no location leave the geometry columns `NA`.

**Filter coverage.** All server-side JSON:API `filter[...]` parameters:
`query` -> `filter[text_search]` (free-text), `type` -> `filter[type]`
(decision-type enum — see `get_assessments_coverage()`, e.g. `"aepo_creation"`),
and `date_range` -> `filter[issued_after]` / `filter[issued_before]` (matched on
the decision issue date, re-checked client-side as a guard). `date_published` is
the registry publication timestamp; `date_decision` is the issue date.

**Documents.** At most one document per decision — the AEPO decision PDF — taken
verbatim from `diavgeia_doc_url`. These are hosted on **Διαύγεια / Diavgeia**
(the Greek government transparency portal) at
`https://diavgeia.gov.gr/luminapi/api/decisions/{ADA}/document.pdf`; the ADA in
the path contains Greek characters, left for the HTTP client to percent-encode
at request time. It lands under the single slug `attachment_urls_aepo` /
`local_path_aepo`, with the deduplicated union at `attachment_urls` /
`local_path`. Records whose `diavgeia_doc_url` is `null` yield zero attachments
(still schema-valid).

```{r}
# AEPO decisions, first few records
get_assessments_gr(limit = 5, download = FALSE)

# New AEPO approvals only
get_assessments_gr(type = "aepo_creation", limit = 5, download = FALSE)
```

## Iceland — `"is"`

* **Portal:** **Skipulagsgátt** (Skipulagsstofnun / National Planning Agency),
  served by an anonymous **GraphQL** API at
  [www.skipulagsgatt.is/graphql](https://www.skipulagsgatt.is/graphql). The
  human-facing Vue SPA lives at
  [www.skipulagsgatt.is](https://www.skipulagsgatt.is/); the canonical `url`
  for a record is `https://www.skipulagsgatt.is/issues/<id>`. This is the
  package's **first GraphQL-backed handler** — the transport is a single HTTP
  `POST .../graphql` with a JSON body `{"query": ..., "variables": ...}`.
* **Status:** supported, **with an important coverage caveat (recent cases
  only)** — see below. Reflected in `get_assessments_coverage()$status` as
  `"supported (GraphQL; cases from ~June 2023 onward)"`.
* **Authentication:** none (anonymous GraphQL).
* **Coverage horizon (read this first).** Skipulagsgátt only carries cases from
  **roughly June 2023 onward** (when the portal went live). The older
  `skipulag.is` / `island.is` "gagnagrunnur umhverfismats" is dead (it now
  redirects to static info) and is **not** crawled. So this handler yields a
  recent-cases-only corpus (~235 environmental-assessment issues at the time of
  writing).
* **Three processes merged.** The portal models every dossier as a unified
  `Issue`; environmental assessment lives in three of its `process`es, selected
  server-side by `processId` and merged into one tibble with an
  `assessment_type` tag: `15` (*matsskylda* screening, EIA-track) and `16`
  (full *Mat á umhverfisáhrifum*, EIA) → `"EIA"`; `501` (*Umhverfismat
  áætlana*, SEA) → `"SEA"`. The `assessment_type` argument
  (`"All"`/`"EIA"`/`"SEA"`) picks the process loop. Because the three share the
  same `Issue.id` space, `document_id` is prefixed `IS-<processId>-<id>` (e.g.
  `IS-16-137`) so they never collide on disk; the finer distinction is kept in
  `process_type` / `native_type`.
* **Enumeration.** The listing is the GraphQL `issueConnection(input, first,
  after)` cursor connection, paginated with `after: endCursor`. `processId` is
  single-valued per query, so the handler loops the selected processes. Detail
  is one `singleIssue(issueId)` call per record, sidecar-first.
* **Status field gotcha.** Record status is read from the plain-string
  `lifecycle` field (e.g. `"done"`, `"process_initialized"`). The handler
  deliberately does **not** request the `issueStatus` enum — it has a
  server-side serialization bug that returns HTTP 500 for the whole query on
  some records.
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.is_throttle_rate")`).
* **Language:** record content is Icelandic (ISO-639-1 `is`).

**Geometry.** `Issue.hasGeography` flags whether a record has a footprint;
`Issue.geographies` returns a GraphQL `FeatureCollection` whose per-feature
`geometry` arrives as a GeoJSON **string** that needs a *second*
`jsonlite::fromJSON` parse. Point / LineString / Polygon are all observed.
Coordinates are already geographic **WGS84 (EPSG:4326 lon/lat)** — *not* the
projected ISN93 grid — so no reprojection happens. When present, the geometry
is written next to the sidecar as `<document_id>.geometry.geojson` (family
FeatureCollection layout, `urn:ogc:def:crs:EPSG::4326`); `geometry_crs` is
`"EPSG:4326"`. Records with `hasGeography = false` / an empty collection leave
the geometry columns `NA`.

**Filter coverage.** Server-side: `query` -> the GraphQL `search` field
(full-text), and `date_range` -> `fromDate` / `toDate` (matched on the
`publishedDate`, re-checked client-side as a guard). `assessment_type` selects
which process(es) to crawl (applied in R, by choosing the process loop).
`date_published` is the `publishedDate`; `date_decision` is the `closedDate`.

**Documents.** Documents live under `phases[].files[]`, each an
`IssuePhaseFile` with a semantic-role `type` and a nested `data` `File`. Only
files with `published == true` are emitted. They are grouped by their role into
lowercase slugs — `almennt` (general; incl. the matsskýrsla / EIA report and
matsáætlun), `vidbrogd` (developer responses), `afgreidsla` (the decision /
álit) — one `attachment_urls_<slug>` / `local_path_<slug>` list-column each,
with the deduplicated union at `attachment_urls` / `local_path`. The download
URL is `https://www.skipulagsgatt.is/<File.path>` (path of the form
`files/<uuid>`), anonymous. PDFs are large (often 10-16 MB) — set
`max_file_size_mb` accordingly. The top-level `Issue.files` collection is
usually empty and is ignored.

```{r}
# Recent EIA cases (screening + full), first few records
get_assessments_is(assessment_type = "EIA", limit = 5, download = FALSE)

# SEA only (umhverfismat áætlana)
get_assessments_is(assessment_type = "SEA", limit = 5, download = FALSE)
```

## Ireland — `"ie"`

* **Portal:** Ireland's **EIA Portal** (gov.ie), an Esri ArcGIS Online map
  app backed by a public **anonymous ArcGIS REST FeatureServer** at
  [services.arcgis.com](https://services.arcgis.com/NzlPQPKn5QF9v2US/arcgis/rest/services/EIA_Location_Point/FeatureServer/0)
  (the `EIA_Location_Point` master layer, ≈5,100 records). The human-facing
  app lives at
  [experience.arcgis.com/.../a1a85d92623147b191dd25a14b2571da](https://experience.arcgis.com/experience/a1a85d92623147b191dd25a14b2571da/).
  This is the package's **first ArcGIS REST-backed handler** — transport is a
  plain `GET .../query?f=json`, pagination is `resultOffset` /
  `resultRecordCount`, and the Esri point geometry is converted to GeoJSON
  in-house.
* **Status:** supported, **with two important caveats** — see below. Reflected
  in `get_assessments_coverage()$status` as
  `"supported (EIA only; portal = notice PDFs, full EIAR off-portal)"`.
* **Authentication:** none (anonymous public layer).
* **EIA only — no SEA.** The portal covers **EIA applications only**; there is
  no SEA register here.
* **Notices only — full EIAR is off-portal.** For each application the portal
  hosts (at most) the statutory **newspaper / public-notice PDF**. The full
  EIAR itself lives **off-portal** on the relevant competent-authority website
  (An Bord Pleanála, the local council, the EPA, …). Those external case pages
  are surfaced as the `url_link_application` / `url_link_secondary` extras
  columns — HTML landing pages on heterogeneous third-party sites, *not* direct
  PDFs — and are deliberately kept **out** of `attachment_urls`. Treat them as a
  discovery target.
* **The `OBJECTID_1` gotcha.** The layer's unique object-id field is
  **`OBJECTID_1`**, *not* the plain `OBJECTID` (which is non-unique on this
  layer and frequently `0`). All attachment lookups and the internal id
  fallback key off `OBJECTID_1`. The stable business key used as `document_id`
  is `Portal_Ref` (e.g. `"2024092"`).
* **Enumeration.** Listing is `GET <layer>/query`, paginated with
  `resultOffset` / `resultRecordCount` (page size 1000), looping while
  `exceededTransferLimit` is true, ordered by `OBJECTID_1 ASC`. Each feature
  carries all its metadata inline in `attributes` (no separate detail endpoint).
  The portal has no per-record permalink, so a unique, deterministic `url` is
  synthesised per record — the record-specific query URL
  `<layer>/query?where=Portal_Ref='<ref>'&outFields=*&f=json`.
* **Dates.** `Date_of_receipt_of_application_` is **epoch milliseconds** →
  `date_published` (divided by 1000). `date_decision` is `NA` (no decision-date
  field on the layer).
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.ie_throttle_rate")`).
* **Language:** English (ISO-639-1 `en`).

**Geometry.** Each record carries an Esri point `geometry` (`{x, y}`) in
**Irish Transverse Mercator (IRENET95 ITM / EPSG:2157)** — requested via
`outSR=2157`. It is converted in-house to a GeoJSON **Point** and, when
`write_sidecar = TRUE`, written next to the sidecar as
`<document_id>.geometry.geojson` (family FeatureCollection layout,
`urn:ogc:def:crs:EPSG::2157`); `geometry_crs` is `"EPSG:2157"`. No reprojection
happens — coordinates stay in EPSG:2157. Records with a null geometry leave the
geometry columns `NA`.

**Filter coverage.** Server-side: `query` -> an
`UPPER(Description__Max__256_character) LIKE '%…%'` predicate (free-text over
the description), `competent_authority` -> equality on `Competent_Authority`,
and `date_range` -> a `Date_of_receipt_of_application_` BETWEEN window (epoch
ms), re-checked client-side against `date_published`.

**Documents.** Portal-hosted attachments are ArcGIS feature attachments — the
statutory newspaper / public-notice PDF. Because the download URL needs both the
`OBJECTID_1` and the per-attachment id
(`<layer>/<OBJECTID_1>/attachments/<id>`), attachment metadata is resolved in a
batched **phase-2** `queryAttachments` call (keyed by `OBJECTID_1`) — even when
`download = FALSE`, this is needed to populate `attachment_urls`. The single
slug is `notice` (`attachment_urls_notice` / `local_path_notice`); the
deduplicated union goes to `attachment_urls` / `local_path`. A record may carry
**0** attachments (still schema-valid). The full EIAR is *not* here — see the
off-portal caveat above.

```{r}
# Recent EIA applications, first few records (portal hosts the notice PDF only)
get_assessments_ie(limit = 5, download = FALSE)

# Wind-themed slice (server-side description LIKE)
get_assessments_ie(query = "wind", limit = 20, download = FALSE)
```

## Slovenia — `"si"`

* **Portal:** the Slovenian government portal **gov.si**, which publishes its
  environmental-assessment registers under the
  [okoljske-presoje](https://www.gov.si/podrocja/okolje-in-prostor/okolje/okoljske-presoje)
  pages. Each register offers a **bulk JSON export** at
  `<list>/export/json/` that returns the whole register as a single JSON array
  (not paginated) — the handler issues one GET per register.
* **Status:** supported (`get_assessments_coverage()$status` = `"supported"`).
* **Authentication:** none (anonymous public exports).
* **Dual-register.** Three registers are merged into one result tibble, tagged
  with `assessment_type` (`"EIA"` / `"SEA"`) and the raw `register` label:
  * EIA screening — `predhodni-postopek` (register `"predhodni-postopek"`,
    `document_id` prefix `PRED-`);
  * SEA — CPVO decisions for **state** spatial plans
    (`odlocitve-…-drzavnih-prostorskih-nacrtov`, register `"cpvo-drzavni"`,
    prefix `CPVO-DRZ-`) and **municipal** spatial plans
    (`odlocitve-…-obcinskih-prostorskih-nacrtov-2`, register `"cpvo-obcinski"`,
    prefix `CPVO-OBC-`).
  An `assessment_type` argument (`"All"` default / `"EIA"` / `"SEA"`) selects
  which register(s) to crawl.
* **Field maps differ per register.** The screening export uses Slovenian keys
  (`Poseg` → title, `Datum objave` → `date_published`, `Naziv` → proponent,
  `Naslov` → `proponent_address`, `Številka zadeve` → `case_number`,
  `Oznaka posega` → `annex_code`); the CPVO exports use `Title`/`Naziv` →
  title, `Datum` → `date_published`, and `Odločitev` → `decision` (also the
  `native_type`). `competent_authority` is the fixed national ministry
  (*Ministrstvo za okolje, podnebje in energijo*). Slovenian values are kept
  verbatim.
* **Attachments.** The bulk-export ids do not map to filenames, so attachments
  are scraped from each record's detail page (sidecar-first): every
  `/assets/seznami/…` link is collected and absolutised against
  `https://www.gov.si`. These go to `attachment_urls` (a flat list, no section
  split). A record may carry **0** attachments (still schema-valid).
* **No geometry.** The registers expose no spatial geometry.
* **Dates.** `date_published` parses the Slovenian `"DD. MM. YYYY"` format;
  `date_decision` is always `NA` (no decision date as a date). `date_range` is
  matched client-side against `date_published`.
* **Scope.** The registers cover roughly **2021 onward**.
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.si_throttle_rate")`).
* **Language:** Slovenian (ISO-639-1 `sl`); no normalisation.

```{r}
# Recent screening + CPVO records (no PDFs downloaded yet)
get_assessments_si(limit = 5, download = FALSE)

# SEA only (both CPVO registers)
get_assessments_si(assessment_type = "SEA", limit = 10, download = FALSE)
```

## Adding a new source

Adding a handler for another national portal is a matter of writing one
`get_assessments_<cc>()` that returns a tibble with the planscanR required
columns (`country`, `source_portal`, `document_id`, `url`, `retrieved_at`,
`attachment_urls`, `local_path`) and wiring it into the dispatcher in
`get_assessments()`. If you have a portal in mind, open an issue describing
what it exposes (sitemap? search? open JSON?) and we can sketch out a handler.
