---
title: "Supported sources"
output:
  rmarkdown::html_vignette:
    pandoc_args: ["--from=markdown+autolink_bare_uris+tex_math_single_backslash+alerts"]
vignette: >
  %\VignetteIndexEntry{Supported sources}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Every chunk in this vignette talks to a portal, so they are shown but
  # not executed at build time.
  eval = FALSE
)
library(planscanR)
```

`planscanR` is a thin layer over a handful of national environmental-assessment
portals. There is no shared API behind the scenes — each country gives us a
different mix of HTML pages, sitemaps, search endpoints, and undocumented JSON
handlers, and each one comes with its own quirks. This vignette describes what
each handler does, where it pulls data from, and what to expect (or watch out
for) when using it.

For the runtime equivalent — the same information in tabular form, including
the valid vocabularies for the search facets each handler accepts — call:

```{r}
get_assessments_coverage()
```

## A note on stability

> [!WARNING]
> None of these portals expose a contractually stable API. Three of the four
> handlers parse HTML detail pages with `rvest`; the fourth uses an
> undocumented JSON service. Any of the operators can change a CSS class,
> rename a field, or slip a login wall in front of an endpoint without
> warning, and the corresponding handler will then break — sometimes
> silently. Treat the results with the scepticism that a scraping pipeline
> deserves, and please file an issue when you spot drift.

> [!CAUTION]
> Because we are often not talking to real APIs, we have no rate-limit
> contract with these servers. The package throttles where we know it
> matters (NL is capped at ~1 request/second, DK at 5), but a careless
> full-register crawl can still put real load on a small government portal.
> Use `limit` and `query` while exploring.

## Netherlands — `"nl"`

* **Portal:** Commissie m.e.r. adviezenregister
  ([commissiemer.nl/adviezen](https://www.commissiemer.nl/adviezen/))
* **Status:** supported — full records and document downloads.
* **Authentication:** none.
* **Coverage:** ~3,600 advisory records.
* **Throttle:** ~1 request/second by default
  (`getOption("planscanR.nl_throttle_rate")`).

URL enumeration is driven by the portal's own sitemap index (`wp-sitemap.xml`,
following the `advice-sitemap*` sub-sitemaps). Per-record metadata is parsed
from each detail page with `rvest` and persisted to a sidecar JSON, so
re-running a slice is essentially free once the cache is warm.

**Filter coverage.** Free-text `query`, `date_range` (matched against
`date_decision`), and `province` (substring match against
`competent_authority`) are honoured. The portal's taxonomy facets — `theme`,
`advice_type`, `status` — are driven by a client-side FacetWP layer that does
not yield to programmatic access; those arguments are accepted for forward
compatibility but currently emit a one-shot warning.

**Documents.** Detail pages group PDFs into two on-page sections, which are
exposed as separate list-columns (`attachment_urls_source` and the
deduplicated `attachment_urls` union). See `?get_assessments_nl` for the
section layout.

```{r}
get_assessments_nl(limit = 20, download = FALSE)
```

## Germany — `"de"`

* **Portal:** UVP-Verbund
  ([uvp-verbund.de](https://www.uvp-verbund.de/))
* **Status:** supported — full records and document downloads.
* **Authentication:** none.
* **Coverage:** ~24,289 records federated across all federal-state authorities
  (the full register).

The portal exposes no sitemap, no OAI-PMH, and no CSW endpoint. Enumeration
goes through the portal's Solr-backed full-text search (`/freitextsuche`).
When no `query` is supplied the handler sends the wildcard `q=*:*`, which
paginates through the full register (~2,429 pages of 10 records each). The
`toggle_procedure=` parameter is explicitly set to an empty value so the
portal does not silently restrict results to currently-running and
last-year-modified procedures.

**Filter coverage.** `query` is passed straight through as the server-side
`q` parameter (real full-text search, not a client-side substring match).
`date_range` matches against the detail-page "Zuletzt geändert" timestamp,
exposed as `date_decision`. `jurisdiction` is a substring match against the
federal-state partner (e.g. `"Bayern"`, `"Baden-Württemberg"`).

**Documents.** Detail pages group documents under open-ended `h4.title-font`
section headings. Each heading becomes its own
`attachment_urls_<slug>` / `local_path_<slug>` column; known headings get a
curated slug (`uvp_bericht`, `berichte`, `entscheidung`, `auslegung`,
`weitere`) and any unknown heading is auto-slugged from its title so a new
section type appears without a code change. See `?get_assessments_de` for the
full layout.

```{r}
get_assessments_de(query = "windenergie", limit = 20, download = FALSE)
```

## Austria — `"at"`

* **Portal:** Umweltbundesamt UVP-DB
  ([secure.umweltbundesamt.at/uvpdb/public](https://secure.umweltbundesamt.at/uvpdb/public))
* **Status:** supported (metadata-only) — documents sit behind a Keycloak
  login wall and are not retrievable by this version.
* **Authentication:** none for metadata.
* **Coverage:** ~500 procedures (the full register is small).

The public HTML pages of the portal sit behind a Keycloak login, but three
JSON service handlers are open: `mapsdata`, `mapsgeom`, and `vorhabenInfo`.
Enumeration is a single `mapsdata` call that returns the full index keyed by
Aktenzahl; per-record detail is one `vorhabenInfo` call per `v2id`. No
pagination, no CSRF, no session required.

**Filter coverage.** `query` is a case-insensitive substring match on
`title` + `summary`. `date_range` is matched against the record's `year`
(treated as a Jan 1 – Dec 31 span) — `date_decision` is always `NA`, because
the anonymous service handlers do not expose a decision or last-modified
timestamp. `jurisdiction` is a substring match against the comma-joined
Austrian federal-state list (so `"Wien"` works, `"Bayern"` will not).

**Documents.** Not available to anonymous callers. Every record returns
`attachment_urls = character(0)`. The `download` argument is accepted for
API symmetry but has no effect. If you need attachments, point
`get_assessments(..., discover = TRUE)` at a web-search backend — see
[`?discover_attachments`].

```{r}
get_assessments_at(query = "Windpark", limit = 20, download = FALSE)
```

## Denmark — `"dk"`

* **Portal:** Danmarks Miljøportal EA-Hub
  ([eahub.miljoeportal.dk](https://eahub.miljoeportal.dk/))
* **Status:** supported (metadata-only) — full metadata and polygon geometry
  are fetched; document downloads are a future addition.
* **Authentication:** none.
* **Coverage:** ~2,700 records (both EIA, "miljøvurdering af projekter", and
  SEA, "miljøvurdering af planer").
* **Throttle:** 5 requests/second by default
  (`getOption("planscanR.dk_throttle_rate")`), because the geometry fetch is
  one tiny GET per record-with-geometry.

EA-Hub is a Vue SPA sitting on a public REST API (Swagger at
`/api/swagger/v1/swagger.json`). One `POST /assessments/search` call returns
the entire register — each row already carries title, year range, status,
authorities, EIA-Directive Annex I/II categories, plan types/categories, and
a `hasGeometry` flag — so no detail call is needed during the scan phase.

**Geometry.** Records with `hasGeometry == TRUE` carry a polygon (typically a
MULTIPOLYGON in EPSG:25832 / ETRS89-UTM32N). When `write_sidecar = TRUE`, the
geometry is fetched from `GET /assessments/{id}/geometry` and saved next to
the sidecar as `<document_id>.geometry.geojson`; the sidecar carries a
`geometry_path` and `geometry_crs` for downstream consumption with `sf`.

**Filter coverage.** `query` is forwarded to the API's server-side `freeText`.
`assessment_type` (`"All"`, `"Plans"`, `"Project"`) is honoured. `date_range`
is matched client-side against each record's `fromYear`/`toYear` (treated as
Jan 1 – Dec 31 spans); `date_decision` is always `NA` because EA-Hub exposes
only year fields, no decision timestamp.

**Documents.** EA-Hub exposes PDFs at public Azure blob URLs reachable via
`GET /assessments/{id}/documents/{docId}/links`, but resolving them costs an
extra HTTP call per document. The current handler is scan + classify only;
a future release will add the download phase.

```{r}
get_assessments_dk(query = "vindmølle", limit = 20, download = FALSE)
```

## Adding a new source

Adding a handler for another national portal is a matter of writing one
`get_assessments_<cc>()` that returns a tibble with the planscanR required
columns (`country`, `source_portal`, `document_id`, `url`, `retrieved_at`,
`attachment_urls`, `local_path`) and wiring it into the dispatcher in
`get_assessments()`. If you have a portal in mind, open an issue describing
what it exposes (sitemap? search? open JSON?) and we can sketch out a handler.
